---
title: "Final_Project_Exploration"
author: "Jack Rossi"
date: "March 19, 2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
#install.packages('nhanesA')
library('nhanesA')
```

## Defining Smoking

SMD030 - "How old {were you/was SP} when {you/s/he} last smoked cigarettes {fairly regularly}?"

SMQ040 {Do you/Does SP} now smoke cigarettes?

```{r}

smok <- nhanes('SMQ_I')
smok$SMQ040c <- smok$SMQ040
smok <- nhanesTranslate('SMQ_I', 'SMQ040', data=smok)
smok <- smok[,c("SEQN","SMD030", "SMQ040", "SMQ040c")]

smok$smok_yn = !is.na(smok$SMD030) 

```

## Defining Bodyweights

WHD020 - "How much {do you/does SP} weigh without clothes or shoes?"

WHD120 - "How much did {you/SP} weigh at age 25?  [If you don't know {your/his/her} exact weight, please make "

WHD130 - "How tall {were you/was SP} at age 25?  [If you don't know {your/his/her} exact height, please make y"

```{r}

wgt <- nhanes('WHQ_I') 
wgt <- wgt[, c("SEQN", "WHD020", "WHD120", "WHD130")]

wgt$bmi <- (wgt$WHD020 * 0.453592) / (wgt$WHD130 * 0.0254)^2

attr(wgt$bmi, 'label') <- "ratio of weight in kg to height in meters"

```

## Alcohol Use

```{r}

alc <- nhanes("ALQ_I")
alc <- nhanesTranslate('ALQ_I','ALQ151', data=alc)
alc <- alc[,c('SEQN', 'ALQ120Q', 'ALQ151')]
```

## Sleep Habits

```{r}
sleep <- nhanes("SLQ_I")
sleep <- nhanesTranslate('SLQ_I', 'SLQ050', data = sleep)
sleep <- sleep[,c("SEQN", "SLD012", "SLQ050")]
attr(sleep$SLQ050, 'label') = "ever told a doctor you had trouble sleeping?"
```

## Diet

```{r}
diet <- nhanes('DBQ_I')
diet <- nhanesTranslate('DBQ_I', 'DBQ700', data = diet)
diet <- diet[,c("SEQN", "DBQ700")]
```

## Nutrients - Vitamin C and Caffeine

```{r}
nutr <- nhanes("DR1TOT_I")

nutr <- nutr[, c("SEQN","DR1TVC","DR1TCAFF")]

```

# Joining all data together

```{r}

full_data <- merge(x = diet ,y = alc ,by="SEQN", all = TRUE)
full_data <- merge(x = full_data, y = nutr, by = "SEQN", all = TRUE)
full_data <- merge(x = full_data, y = sleep, by = "SEQN", all = TRUE)
full_data <- merge(x = full_data, y = smok, by = "SEQN", all = TRUE)
full_data <- merge(x = full_data, y = wgt, by = "SEQN", all = TRUE)

sum(complete.cases(full_data))

head(rowSums(is.na(full_data)))
colSums(is.na(full_data))

save(full_data, file = "Cov.RData")

```

## Defining Hypertension (High Blood Pressure)

BPQ035 - "How old {were you/was SP} when {you were/he/she was} first told that {you/he/she} had hypertension o"

BPQ020 - "{Have you/Has SP} ever been told by a doctor or other health professional that {you/s/he} had hypert"

BPQ030 - "{Were you/Was SP} told on 2 or more different visits that {you/s/he} had hypertension, also called h[igh blood pressure]"


BPQ080 - "{Have you/Has SP} ever been told by a doctor or other health professional that {your/his/her} blood [cholesterol is high]?"


```{r}

bp <- nhanes("BPQ_I")
bp$BPQ020c <- bp$BPQ020
bp$BPQ030c <- bp$BPQ030
bp$BPQ080c <- bp$BPQ080

bp <- nhanesTranslate('BPQ_I', c('BPQ080', "BPQ030", "BPQ020"), data=bp)

bp <- bp[, c("SEQN", "BPD035", 'BPQ080', "BPQ030", "BPQ020", 'BPQ080c', "BPQ030c", "BPQ020c")]

```
