---
title: "Biomed_final"
author: "Tianran Zhang"
date: "4/6/2020"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
library(table1)
library(ggpubr)
library(ggmosaic)
# Effect displays for GLM and more
library(effects)
library(ROCR)
library(pROC)

library(ordinal)
library(MASS)

library(nnet)
```

# Load the data

```{r}
load("./covariates.RData")
load("./Disease.RData")

dat <- merge(x = full_data, y = disease[c("SEQN", "bp", "bi", "cd")], 
             by = "SEQN", all = TRUE)

dat <- dat[complete.cases(dat), ]
dat$ncd <- dat$bp == 1 | dat$bi == 1 | dat$cd == 1

dat <- dat[dat$ALQ151 != "Don't know", ]
dat <- dat[dat$DBQ700 != "Don't know", ]
dat$ALQ151 <- factor(dat$ALQ151)
dat$SLQ050 <- factor(dat$SLQ050)
dat$DBQ700 <- factor(dat$DBQ700)

dat$DBQ700 = relevel(dat$DBQ700, ref = "Excellent")
dat$ALQ151 = relevel(dat$ALQ151, ref = "Yes")
dat$SLQ050 = relevel(dat$SLQ050, ref = "Yes")

head(dat)
```


# Data Summary
## Numerically Summary
```{r}
dat$ncd_p <- factor(dat$ncd, levels = c("TRUE", "FALSE", "P"), labels = c("TRUE", "FALSE", "P-value"))

rndr <- function(x, name, ...){
  if (length(x) == 0){
    y <- dat[[name]]
    s <- rep("", length(render.default(x = y, name = name)))
    if (is.numeric(y)){
      p <- t.test(y ~ dat$ncd_p)$p.value
    } else{
      p <- chisq.test(table(y, droplevels(dat$ncd_p)))$p.value
    }
    s[2] <- sub("<", "&lt;", format.pval(p, digits = 3, eps = 0.001))
    s
  } else{
    render.default(x = x, name = name, ...)
  }
}

rndr.strat <- function(label, n, ...){
  ifelse(n == 0, label, render.strat.default(label, n, ...))
}

table1(~ DBQ700 + ALQ120Q + ALQ151 + DR1TVC + DR1TCAFF + SLD012 + SLQ050 + 
         SMD030 + WHD020 + bmi| ncd_p, data = dat, droplevels = F, 
       render = rndr, render.strat = rndr.strat, overall = F)

```



## Graphical Summary
**Continuous Variable**: ALQ120Q, DR1TVC, DR1TCAFF, SLD012, SMD030, WHD020, bmi   
(Draw Boxplots for the continuous variables)
```{r}
ggboxplot(data = dat, x = "ncd", y = "ALQ120Q", add = "jitter", color = "ncd") + 
  theme(legend.position = "none") + 
  xlab("Having one or more diseases") +
  ylab("How often drink alcohol over past 12 mos") +
  stat_compare_means(method = "anova", label.x.npc = "center", label.y.npc = "top")
ggboxplot(data = dat, x = "ncd", y = "DR1TVC", add = "jitter", color = "ncd")+
  theme(legend.position = "none") + 
  xlab("Having one or more diseases") +
  ylab("Vitamin C (mg)") +
  stat_compare_means(method = "anova", label.x.npc = "center", label.y.npc = "top")
ggboxplot(data = dat, x = "ncd", y = "DR1TCAFF", add = "jitter", color = "ncd")+
  theme(legend.position = "none") +
  xlab("Having one or more diseases") +
  ylab("Caffeine (mg)") +
  stat_compare_means(method = "anova", label.x.npc = "center", label.y.npc = "top")
ggboxplot(data = dat, x = "ncd", y = "SLD012", add = "jitter", color = "ncd")+
  theme(legend.position = "none") + 
  xlab("Having one or more diseases") +
  ylab("Sleep hours") +
  stat_compare_means(method = "anova", label.x.npc = "center", label.y.npc = "top")
ggboxplot(data = dat, x = "ncd", y = "SMD030", add = "jitter", color = "ncd")+
  theme(legend.position = "none") + 
  xlab("Having one or more diseases") +
  ylab("Age started smoking cigarettes regularly") +
  stat_compare_means(method = "anova", label.x.npc = "center", label.y.npc = "top")
ggboxplot(data = dat, x = "ncd", y = "WHD020", add = "jitter", color = "ncd")+
  theme(legend.position = "none") + 
  xlab("Having one or more diseases") +
  ylab("Current self-reported weight (pounds)") +
  stat_compare_means(method = "anova", label.x.npc = "center", label.y.npc = "top")
ggboxplot(data = dat, x = "ncd", y = "bmi", add = "jitter", color = "ncd")+
  theme(legend.position = "none") + 
  xlab("Having one or more diseases") +
  ylab("BMI") +
  stat_compare_means(method = "anova", label.x.npc = "center", label.y.npc = "top")
```

**Discrete variable** DBQ700, ALQ151, SLQ050, smok_yn
(Draw Mosaic plots for discrete variables)
```{r,warning=FALSE}
pval <- function(p){
  if (p < 0.001) {
    return("< 0.001")
    } else {return(round(p, 3))}
}

ggplot(data = dat) + 
  geom_mosaic(aes(x = product(DBQ700, ncd), fill = DBQ700)) +
  xlab(label = "Having one or more diseases") +
  ylab(label = "Diet") +
  theme(legend.position = "none") +
  geom_text(x = .8, y = .8, 
            label = paste0("Chisq p ", 
                           pval(chisq.test(dat$ncd, dat$DBQ700, correct = F)$p.value)))
ggplot(data = dat) + 
  geom_mosaic(aes(x = product(ALQ151, ncd), fill = ALQ151)) +
  xlab(label = "Having one or more diseases") +
  ylab(label = "Alcohol") +
  theme(legend.position = "none") +
  geom_text(x = .8, y = .8, 
            label = paste0("Chisq p ", 
                           pval(chisq.test(dat$ncd, dat$ALQ151, correct = F)$p.value)))
ggplot(data = dat) + 
  geom_mosaic(aes(x = product(SLQ050, ncd), fill = SLQ050)) +
  xlab(label = "Having one or more diseases") +
  ylab(label = "Having trouble sleeping") +
  theme(legend.position = "none") +
  geom_text(x = .8, y = .8, 
            label = paste0("Chisq p ", 
                           pval(chisq.test(dat$ncd, dat$SLQ050, correct = F)$p.value)))
```


# Test 
**Continuous Variable**: ALQ120Q, DR1TVC, DR1TCAFF, SLD012, SMD030, WHD020, bmi      
(Use two sample t-test with two-sided 0.05 significance level to compare the mean level of each variable between two NCDs outcomes)  
```{r,warning=FALSE}
t.test(dat$ncd,dat$ALQ120Q)
t.test(dat$ncd,dat$DR1TVC)
t.test(dat$ncd,dat$DR1TCAFF)
t.test(dat$ncd,dat$SLD012)
t.test(dat$ncd,dat$SMD030)
t.test(dat$ncd,dat$WHD020)
t.test(dat$ncd,dat$bmi)
```

**Discrete variable** DBQ700, ALQ151, SLQ050, smok_yn
(Use Chi-square tests to compare the proportions of diagnosed with NCDs for each discrete variable)    

```{r,warning=FALSE}
chisq.test(dat$ncd,dat$DBQ700)
chisq.test(dat$ncd,dat$ALQ151)
chisq.test(dat$ncd,dat$SLQ050)
```

# Generalized Linear Regression  
## Univariate Regression 
```{r,message=FALSE}
# Univariate logistic regression
colnames(dat) <- c("SEQN","Diet","Alcohol_freq","Alcohol",
                   "Vitamin_C","Caffeine","Sleep","Trouble_sleep",
                   "Smoking_age","smok_yn","Weight","BMI","bp","bi","cd","ncd","ncd_p")
univ = function(var)
{
  form = formula(paste0("ncd ~ ", var))
  logit = glm(form, data = dat, family = "binomial")
  
  p = coef(summary(logit))[,4]
  
  OR = data.frame(round(exp(cbind(OR = coef(logit), confint(logit))), 3),
                  p_value = format.pval(p, digits = 3, eps = 0.001))[-1,]
  colnames(OR) <- c("OR", "2.5%", "97.5%", "p_value")
  return(OR)
}

varlist = colnames(dat)[!(colnames(dat) %in% c("SEQN", "bp", "bi", "cd","ncd","smok_yn","ncd_p"))]

table_uni <- do.call(rbind.data.frame, lapply(varlist, univ))
knitr::kable(table_uni)
```

# Regression
## Full Model
```{r}
form <- formula(paste0("ncd ~" , paste(varlist, collapse = "+")))
full <- glm(form, data = dat, family = "binomial")

final <- step(full, direction = "both", trace = 0)
#summary(final)

p <- coef(summary(final))[,4]
sum_final <- data.frame(round(exp(cbind(OR = coef(final), confint(final))), 4),
           p_value = format.pval(p, digits = 3, eps = 0.001))
colnames(sum_final) <- c("OR", "2.5%", "97.5%", "P-val")
rownames(sum_final)[1] <- "DietExcellent"
sum_final[1,]<-c("Ref","","","")
knitr::kable(sum_final) 
```

# effect plot
```{r,fig.height=10,fig.width=15}
## Effect displays
plot(allEffects(final))
```

# ROC plot
```{r}
rocplot <- function(truth, pred, ...){
  predob = prediction(pred, truth)
  perf = performance(predob, "tpr", "fpr")
  plot(perf, ...) 
  area = auc(truth, pred)
  area = format(round(area, 4), nsmall = 4)
  text(x = .8, y = .1, labels = paste("AUC = ", area))
  segments(x0 = 0, y0 = 0, x1 = 1, y1 = 1, col = "gray", lty = 2)
}

rocplot(dat$ncd, final$fitted.values)
```





# Ordinary

## Data Summary
### Numerically Summary

```{r}
dat2 <- dat %>%
  mutate(ncd = bp + bi + cd)
dat2$ncd <- factor(dat2$ncd)
dat2 <- dat2 %>%
  dplyr::select(-smok_yn, -Weight)
head(dat2)

```

```{r}
dat2$ncd_p <- factor(dat2$ncd, levels = 0:4, labels = c(0:3, "P-value"))

rndr2 <- function(x, name, ...){
  if (length(x) == 0){
    y <- dat2[[name]]
    s <- rep("", length(render.default(x = y, name = name)))
    if (is.numeric(y)){
      p <- kruskal.test(y ~ dat2$ncd_p)$p.value
    } else{
      p <- chisq.test(table(y, droplevels(dat2$ncd_p)))$p.value
    }
    s[2] <- sub("<", "&lt;", format.pval(p, digits = 3, eps = 0.001))
    s
  } else{
    render.default(x = x, name = name, ...)
  }
}

rndr.strat <- function(label, n, ...){
  ifelse(n == 0, label, render.strat.default(label, n, ...))
}

table1(~ Diet+Alcohol+Vitamin_C+Caffeine+Sleep+Trouble_sleep+
         Smoking_age + BMI| ncd_p, data = dat2, droplevels = F, 
       render = rndr2, render.strat = rndr.strat, overall = F)

```


### Graphical Summary
**Continuous Variable**: Alcohol_freq, Vitamin_C, Caffeine, Sleep, Smoking_age, BMI
(Draw Boxplots for the continuous variables)
```{r}
ggboxplot(data = dat2, x = "ncd", y = "Alcohol_freq", add = "jitter", color = "ncd") + 
  theme(legend.position = "none") + 
  xlab("Having one or more diseases") +
  ylab("How often drink alcohol over past 12 mos") +
  stat_compare_means(method = "anova", label.x.npc = "center", label.y.npc = "top")
ggboxplot(data = dat2, x = "ncd", y = "Vitamin_C", add = "jitter", color = "ncd")+
  theme(legend.position = "none") + 
  xlab("Having one or more diseases") +
  ylab("Vitamin C (mg)") +
  stat_compare_means(method = "anova", label.x.npc = "center", label.y.npc = "top")
ggboxplot(data = dat2, x = "ncd", y = "Caffeine", add = "jitter", color = "ncd")+
  theme(legend.position = "none") +
  xlab("Having one or more diseases") +
  ylab("Caffeine (mg)") +
  stat_compare_means(method = "anova", label.x.npc = "center", label.y.npc = "top")
ggboxplot(data = dat2, x = "ncd", y = "Sleep", add = "jitter", color = "ncd")+
  theme(legend.position = "none") + 
  xlab("Having one or more diseases") +
  ylab("Sleep hours") +
  stat_compare_means(method = "anova", label.x.npc = "center", label.y.npc = "top")
ggboxplot(data = dat2, x = "ncd", y = "Smoking_age", add = "jitter", color = "ncd")+
  theme(legend.position = "none") + 
  xlab("Having one or more diseases") +
  ylab("Age started smoking cigarettes regularly") +
  stat_compare_means(method = "anova", label.x.npc = "center", label.y.npc = "top")
# ggboxplot(data = dat2, x = "ncd", y = "Weight", add = "jitter", color = "ncd")+
#   theme(legend.position = "none") + 
#   xlab("Having one or more diseases") +
#   ylab("Current self-reported weight (pounds)") +
#   stat_compare_means(method = "anova", label.x.npc = "center", label.y.npc = "top")
ggboxplot(data = dat2, x = "ncd", y = "BMI", add = "jitter", color = "ncd")+
  theme(legend.position = "none") + 
  xlab("Having one or more diseases") +
  ylab("BMI") +
  stat_compare_means(method = "anova", label.x.npc = "center", label.y.npc = "top")
```

**Discrete variable** Diet, Alcohol, Trouble_sleep
(Draw Mosaic plots for discrete variables)
```{r,warning=FALSE}
pval <- function(p){
  if (p < 0.001) {
    return("< 0.001")
    } else {return(round(p, 3))}
}

ggplot(data = dat2) + 
  geom_mosaic(aes(x = product(Diet, ncd), fill = Diet)) +
  xlab(label = "Having one or more diseases") +
  ylab(label = "Diet") +
  theme(legend.position = "none") +
  geom_text(x = .8, y = .8, 
            label = paste0("Chisq p ", 
                           pval(chisq.test(dat$ncd, dat$DBQ700, correct = F)$p.value)))
ggplot(data = dat2) + 
  geom_mosaic(aes(x = product(Alcohol, ncd), fill = Alcohol)) +
  xlab(label = "Having one or more diseases") +
  ylab(label = "Alcohol") +
  theme(legend.position = "none") +
  geom_text(x = .8, y = .8, 
            label = paste0("Chisq p ", 
                           pval(chisq.test(dat$ncd, dat$ALQ151, correct = F)$p.value)))
ggplot(data = dat2) + 
  geom_mosaic(aes(x = product(Trouble_sleep, ncd), fill = Trouble_sleep)) +
  xlab(label = "Having one or more diseases") +
  ylab(label = "Having trouble sleeping") +
  theme(legend.position = "none") +
  geom_text(x = .8, y = .8, 
            label = paste0("Chisq p ", 
                           pval(chisq.test(dat$ncd, dat$SLQ050, correct = F)$p.value)))
```


## Test 
**Continuous Variable**: ALQ120Q, DR1TVC, DR1TCAFF, SLD012, SMD030, WHD020, bmi      
(Use two sample t-test with two-sided 0.05 significance level to compare the mean level of each variable between two NCDs outcomes)  
```{r,warning=FALSE}
kruskal.test(dat2$ncd,dat2$Alcohol_freq)
kruskal.test(dat2$ncd,dat2$Vitamin_C)
kruskal.test(dat2$ncd,dat2$Caffeine)
kruskal.test(dat2$ncd,dat2$Sleep)
kruskal.test(dat2$ncd,dat2$Smoking_age)
# kruskal.test(dat2$ncd,dat2$Weight)
kruskal.test(dat2$ncd,dat2$BMI)
```

**Discrete variable** DBQ700, ALQ151, SLQ050, smok_yn
(Use Chi-square tests to compare the proportions of diagnosed with NCDs for each discrete variable)    

```{r,warning=FALSE}
chisq.test(dat2$ncd,dat2$Diet)
chisq.test(dat2$ncd,dat2$Alcohol)
chisq.test(dat2$ncd,dat2$Trouble_sleep)
```

## For each covariate fit the univariate logistic model.
### Check the proportional odds assumption
```{r}
# Check the Proportional odds assumption for each variable

varlist2 <- colnames(dat2)[!(colnames(dat2) %in% c("SEQN", "bp", "bi", "cd", "ncd", "ncd_p"))]
prop_test <- data.frame()
for (var in varlist2){
  form = formula(paste0("ncd ~ ", var))
  logit = clm(form, data = dat2)
  a = scale_test(logit)$`Pr(>Chi)`[2]
  b = nominal_test(logit)$`Pr(>Chi)`[2]
  OR = data.frame(round(exp(cbind(OR = coef(logit)[4], confint(logit))), 3))
  p1 <- cbind(OR, p_val = coef(summary(logit))[-1:-3, 4] < 0.05, 
              satisfy_normality = a > 0.05,
              satisfy_effect = b > 0.05)
  prop_test <- rbind(prop_test, p1)
}
prop_test
```

All other variables are violated the PH assumption.


## Univariate
```{r}
comby <- function(y, j){
  ans <- data.frame()
  for (i in 1:3){
    ans <- rbind(ans, y[,,i][j, ])
  }
  ans
}

comb <- function(x, y){
  ans <- data.frame()
  i = 3
  for (i in 2:ncol(x)){
    ans1 <- cbind(x[, i], comby(y, i))
    colnames(ans1) <- c("OR", "2.5%", "97.5%")
    rownames(ans1) <- paste0(colnames(x)[i], 1:3)
    ans <- rbind(ans, ans1)
  }
  ans
}

univ <- function(var){
  form = formula(paste0("ncd ~ ", var))
  multi = multinom(form, data = dat2)
  comb(exp(coef(multi)), exp(confint(multi)))
}

t<- round(do.call(rbind.data.frame, lapply(varlist2, univ)), 3)
knitr::kable(t)
```


```{r}
form2 <- formula(paste0("ncd ~" , paste(varlist2, collapse = "+")))
full2 <- clm(ncd ~ Diet + Alcohol + Vitamin_C + Caffeine + 
    Sleep + Trouble_sleep + Smoking_age + BMI, data = dat2)

final2 <- step(full2, direction = "both", trace = 0)
sum_final2 <- round(cbind(exp(coef(final2)[4:12]), exp(confint(final2))),3)
colnames(sum_final2)[1] <- "OR"
DietExcellent<-c("Ref","","")
sum_final2<-rbind(DietExcellent,sum_final2)
knitr::kable(sum_final2)
# FYI, Interaction: Not significant
```


# effect plot
```{r,fig.height=10,fig.width=15}
## Effect displays
plot(allEffects(full2))
```

# ROC plot
```{r}
multiclass.roc(dat2$ncd, final2$fitted.values)
```

